`gosset`: An R package for analysis and synthesis of ranking data in agricultural

### 2022-07-08 inst/paper/paper.Rmd
Update paper.Rmd
minor edits

The analysis of ranking data requires the use of appropriate statistical models such as the Plackett-Luce model [@luce_individual_1959; @Plackett] or the Bradley-Terry model [@bradley1952rank]. Functionality for fitting these models are available in R with the packages BradleyTerry2 [@BradleyTerry2] and PlackettLuce [@Turner2020] respectively. However, extended functionality was required for the entire data science workflow, which usually includes: (1) Data preparation and cleaning, (2) modelling and validation, and (3) results presentation. For (1) gosset provides functions for converting and preparing data into a ranking or pairwise format required by the packages PlackettLuce and BradleyTerry2 respectively. For (2), gosset provides functions for model selection and validation using cross-validation. In the case of (3), enhanced functionality for plotting model results is provided by the gosset package.

@@

To consider the effect of climate factors on yield, we use agro-climatic covariates to fit a Plackett-Luce tree. For simplicity, we use the total rainfall (Rtotal) derived from CHIRPS data [@Funk2015], obtained in R using the R package chirps [@chirps]. Additional covariates can be used in a Plackett-Luce tree, for example using temperature data from R package ag5Tools [@ag5tools] or nasapower [@nasapower].

>>>>

The analysis of ranking data requires the use of appropriate statistical models such as the Plackett-Luce model [@luce_individual_1959; @Plackett] or the Bradley-Terry model [@bradley1952rank]. Functionality for fitting these models is available in R with the packages BradleyTerry2 [@BradleyTerry2] and PlackettLuce [@Turner2020] respectively. However, extended functionality was required for the entire data science workflow, which usually includes: (1) Data preparation and cleaning, (2) modelling and validation, and (3) results presentation. For (1) gosset provides functions for converting and preparing data into a ranking or pairwise format required by the packages PlackettLuce and BradleyTerry2 respectively. For (2), gosset provides functions for model selection and validation using cross-validation. In the case of (3), enhanced functionality for plotting model results is provided by the gosset package.

@@

To consider the effect of climate factors on yield, we use agro-climatic covariates to fit a Plackett-Luce tree. For simplicity, we use the total rainfall (Rtotal) derived from CHIRPS data [@Funk2015], obtained using the R package chirps [@chirps]. Additional covariates can be used in a Plackett-Luce tree, for example using temperature data from R package ag5Tools [@ag5tools] or nasapower [@nasapower].
